Granular synthesis for instrument makers

### 2022-05-03 paper.tex
Updated paper. Added system diagram.

\documentclass[12pt, IEEEtran]{article}
\usepackage[utf8]{inputenc}
\title{Granular synthesis in practice}
\author{Daniël Kamp}
\date{April 2022}

@@

\begin{abstract}
This writing aims to provide insight into various forms of granular synthesis. It lists existing implementations, explains the inner workings of such systems, and provides insight into human interfaces to use this technique.
\end{abstract}

@@

\subsection{A brief history}
In this subsection I give a brief overview of the history of this synthesis technique.

\subsection{General system overview}
In this section, I want to list the general parts that make up a granular synthesis system. [insert block diagram]
\subsubsection{Sound source}
Here I write about samples, chopping, et cetera.
\subsubsection{Windowing and envelope}
Here I write about windowing, talk about some problems, and provide different envelope shapes and formulas.
\subsubsection{Various parameters}
Here I list various parameters that might be implemented along a granular system.
\begin{itemize}
\item{Stereo panning}
\item{Spread}
\item{something}
\end{itemize}

\subsection{Existing instruments}
In this section, I list a number of instruments that I believe to be important in the world of granular synthesis.

\subsubsection{Mutable Instruments Clouds}
Here I describe MI Clouds.

@@

\subsubsection{Output Portal}
Here I describe Portal

\subsubsection{Instruo Arbhar}
Here I describe Arbhar

\subsubsection{Waldorf Quantum/Iridium}
Here I describe Quantum

@@

\subsection{Implementations}
This subsection lists different implementations.

@@

\section{How is granular synthesis used?}
This section looks at the human aspect of granular synthesis: what sounds does it produce, how is it controlled, and how can one play an instrument that uses this technique.

>>>>

\documentclass[10pt, twocolumn]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\title{Granular synthesis for instrument makers}
\author{	
	\IEEEauthorblockN{Daniël Kamp\\}
    \IEEEauthorblockA{HKU University of the Arts Utrecht
    \\daniel.kamp@student.hku.nl}
    }
\date{April 2022}

@@

\begin{abstract}
This writing aims to provide insight into various forms of granular synthesis. It lists some existing implementations, explains the inner workings of such a system, and provides insight into human interfaces to use this technique.
\end{abstract}

@@

\subsection{A brief history}
Granular synthesis as it's known today is a form of digital sound synthesis, which is based on the processing of small slices of sampled audio known as "grains". The underlying view on sound perception was first proposed by British physicist Dennis Gabor in an article titled "Acoustical Quanta and the Theory of Hearing", published in 1947. In this writing, Gabor theorized

\subsection{General system overview}
This section describes the general anatomy of a granular synthesizer, and gives a broad overview of the building blocks that make up such a system. It's expected that the reader has a basic knowledge of the workings of digital audio, and is familiar with the concept of sampling.

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{GranularDiagram.png}
	\caption{A simplified overview of a granular synthesis system.}
	\label{fig:block_diagram}
\end{figure}

\subsubsection{Sound source}
Here I write about samples, chopping, et cetera.
\subsubsection{Slicing}
Here I write about slicing (transients etc)
\subsubsection{Windowing and envelope}
Here I write about windowing, talk about some problems, and provide different envelope shapes and formulas.
\subsubsection{Various parameters}
Here I list various parameters that might be implemented along a granular system.
\begin{itemize}
\item{Stereo panning}
\item{Spread}
\item{something}
\end{itemize}

\subsection{Existing instruments}
This section lists a few instruments and effects that use granular synthesis. Author's note: I selected open-source instruments of which the source code is freely available to the public. This way, I can give an in-depth comparison of the inner workings of these instruments.

\subsubsection{Mutable Instruments' Clouds}
Clouds is a Eurorack module developed and produced by Émilie Gilet (Mutable Instruments). Both the hardware files (schematics, CAD-files) and firmware source code are open-source and publicly available.
What?
- Eurorack module
- Open-source hardware and firmware
- Granular effect with built-in sound source
How?
- Clouds triggers one grain at a time
- Clouds uses a predetermined amount of grains (64 max)
- Clouds uses Look Up Tables for almost everything (filter, windowing, etc)

@@

\subsubsection{Argotlunar}
Argotlunar is a realtime granular VST/AudioUnit plugin created by Michael Ourednik. The source code of this software is open source and available via Github.

@@

\subsection{Implementations}
This subsection lists different implementations.
\subsubsection{Sound sourcing and selection}
This part lists different methods of sound sourcing and sample selection:
- Microphone or audio input
- Sample-based granular synthesis
For sample selection:
- Randomized selection
- Transient selection
\subsubsection{Trigger algorithms}
Here I write about different trigger algorithms: when does a grain get triggered, what conditions exist, and how can this be controlled in a musical manner?
\subsubsection{Windowing algorithms and approaches}
Here I write about different approaches to windowing: look-up tables, real-time calculation, pre-calculation, etc.
Note: Mutable Instruments uses LUTs, I want to use runtime calculation (insert tested efficiency of both)

@@

\section{How is granular synthesis used?}
This section looks at the human aspect of granular synthesis: what sounds does it produce, how is it controlled, and how can one play an instrument that uses this technique. KEEP THIS PART SHORT
